---
title: "Project 1"
output: html_notebook
---

## Initial Setup

```{r}
#Install and activate packages
pkgs <- c(
  "quantmod", "tidyverse", "lubridate",
  "tseries", "forecast",
  "dynlm", "lmtest", "sandwich", "zoo"
)

installed <- rownames(installed.packages())

for (p in pkgs) {
  if (!p %in% installed) {
    install.packages(p, dependencies = TRUE)
  }
  library(p, character.only = TRUE)
}
```

## Getting Our Data From FRED

```{r}
#We will are focusing on 2000 to present
start_date <- as.Date("2000-01-01")
end_date   <- Sys.Date()

#Data we only care about
symbols <- c("FEDFUNDS", "UNRATE", "CPIAUCSL")

#Get data from FRED via package
getSymbols(Symbols = symbols, src = "FRED", from = start_date, to = end_date)

# Merge all series safely by date + omit NAs
data_xts <- na.omit(merge(FEDFUNDS, UNRATE, CPIAUCSL))

# Combine into one monthly data frame
df <- tibble(
  date     = as.Date(index(data_xts)),
  fedfunds = as.numeric(data_xts$FEDFUNDS),
  unrate   = as.numeric(data_xts$UNRATE),
  cpi      = as.numeric(data_xts$CPIAUCSL)
)

head(df)
```

## Creating a Time Series of our Data

```{r}
#Create plot of all 3 in a row to compare
df_long <- df %>%
  pivot_longer(-date, names_to = "series", values_to = "value")

ggplot(df_long, aes(x = date, y = value)) +
  geom_line() +
  facet_wrap(~series, scales = "free_y", ncol = 1) +
  labs(title = "Time Series Plots (FRED)", x = NULL, y = NULL)
```

## Checking for Seasonality

```{r}
#Check for seasonality
ggseasonplot(ts(df$unrate, frequency = 12,
                start = c(year(min(df$date)), month(min(df$date)))),
             year.labels = TRUE) +
  ggtitle("Seasonal Plot: Unemployment Rate")
```

## Create Separate Dataframe for our Model

```{r}
#Create a new DF to create our DL model. Differencing the values for UNRATE and FEDFUNDS
df2 <- df %>%
  arrange(date) %>%
  mutate(
    d_unrate   = unrate - lag(unrate),
    d_fedfunds = fedfunds - lag(fedfunds),
    infl       = 100 * (log(cpi) - log(lag(cpi)))
  ) %>%
  drop_na()

summary(df2)
```

## ADF Tests

```{r}
adf.test(df$unrate)        
adf.test(df2$d_unrate)     #differenced

adf.test(df$fedfunds)      
adf.test(df2$d_fedfunds)   #differenced

adf.test(df2$infl)
```

## ACF/PACF plots

```{r}
#Create ACF/PACF plots of variables
par(mfrow = c(2,2))
acf(df$unrate, main = "ACF: UNRATE (level)")
pacf(df$unrate, main = "PACF: UNRATE (level)")
acf(df2$d_unrate, main = "ACF: d_UNRATE")
pacf(df2$d_unrate, main = "PACF: d_UNRATE")
par(mfrow = c(1,1))
```

## Creating a Lagged Dataframe

```{r}
#Creating a Lagged Dataframe. Using 12 lags for econmoic purposes
make_lags <- function(x, K, prefix = "x") {
  out <- map_dfc(0:K, ~dplyr::lag(x, .x))
  names(out) <- paste0(prefix, "_L", 0:K)
  out
}

K <- 12

Xlags <- make_lags(df2$d_fedfunds, K = K, prefix = "dff")
df_m <- bind_cols(df2, Xlags) %>%
  drop_na()
```

## Building our Finite Distributed Lag (FDL) Model

```{r}
# Build formula: d_unrate ~ dff_L0 + ... + dff_LK + infl
lag_terms <- paste0("dff_L", 0:K, collapse = " + ")
fml <- as.formula(paste("d_unrate ~", lag_terms, "+ infl"))

dl_fit <- lm(fml, data = df_m)
summary(dl_fit)
```

## AIC/BIC Test for Lag Selection

```{r}
#Lag selection using AIC/BIC
fit_by_K <- function(K) {
  Xlags <- make_lags(df2$d_fedfunds, K = K, prefix = "dff")
  dm <- bind_cols(df2, Xlags) %>% drop_na()
  lag_terms <- paste0("dff_L", 0:K, collapse = " + ")
  fml <- as.formula(paste("d_unrate ~", lag_terms, "+ infl"))
  fit <- lm(fml, data = dm)
  tibble(K = K, AIC = AIC(fit), BIC = BIC(fit))
}

ic_tbl <- map_dfr(0:18, fit_by_K)
ic_tbl %>% arrange(BIC) %>% print(n = 10)

ggplot(ic_tbl, aes(x = K)) +
  geom_line(aes(y = AIC)) +
  geom_line(aes(y = BIC)) +
  labs(title = "Lag Length Selection via AIC/BIC",
       y = "Information Criterion", x = "Max Lag K")
```

## Newey-West Estimator

```{r}
#Newey-West for creating robust standard errors
coeftest(dl_fit, vcov. = NeweyWest(dl_fit, lag = 6, prewhite = FALSE))
```

## Plot distributed lag effects

```{r}
#Plot distributed lag effects
b <- coef(dl_fit)
lag_b <- tibble(
  lag = 0:K,
  beta = b[paste0("dff_L", 0:K)]
)

ggplot(lag_b, aes(x = lag, y = beta)) +
  geom_line() +
  geom_point() +
  labs(title = "Distributed Lag Effects: ΔFedFunds on ΔUnemployment",
       x = "Lag (months)", y = "Coefficient")
```

## Plot ACF/PACF Residuals

```{r}
#Plot ACF/PACF Residuals
par(mfrow = c(1,2))
acf(residuals(dl_fit), main = "ACF: DL Residuals")
pacf(residuals(dl_fit), main = "PACF: DL Residuals")
par(mfrow = c(1,1))
```

## Durbin-Watson Test

```{r}
#Durbin-Watson Test
dwtest(dl_fit) 
```

## Breusch-Godfrey

```{r}
#Breusch-Godfrey for higher-order serial correlation
bgtest(dl_fit, order = 12)
```

## Breusch-Pagan Test

```{r}
#Breusch-Pagan Test
bptest(dl_fit)  
```
